### 2023-09-03 sections/2-methods.tex
table and knn complexity

We group the Sieve-based algorithms together, as their complexity analyses are similar.

The complexity of the Sieve-based algorithms relies on the assumption that when the cardinalities of clusters is sufficiently close to $k$, the distribution of clusters will mimic the distribution of points in the dataset. 
From this assumption, we can infer that number of clusters which overlap with the query ball is a function of the local fractal dimension near the query point.
In particular, we assume that the number of clusters which overlap the query ball is bounded above by 2$d$, where $d$ is the local fractal dimension of the dataset near the query point, as at each end of each of the $d$ axes, we could have a cluster which overlaps the query ball.

In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log_2{\frac{n}{k}}$ iterations of Sieve to reduce the cardinality of clusters to at most $k$.

>>>>

We group the Sieve-based algorithms together, as their complexity analyses are similar.
The complexity of Sieve and Sieve with Separate Centers is dominated by the expected 
cost of Quicksort-inspired partition algorithm we use to calculate thresholds. This method is linear in the size of the input list $Q$, so we must estimate the size of $Q$ on each iteration.

We rely on the assumption that when the cardinalities of clusters is within a low factor of $k$, say $2k$, the distribution of clusters will mimic the distribution of points in the dataset. 
From this assumption, we can infer that number of clusters which overlap with the query ball is a function of the local fractal dimension near the query point.
In particular, we assume that the number of clusters which overlap the query ball is bounded above by 2$d$, where $d$ is the local fractal dimension of the dataset near the query point, as we could have a cluster overlapping the query ball at each end of each of the $d$ axes.

In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log_2{\frac{n}{2k}}$ iterations of Sieve to partition clusters enough that their cardinalities are at most $2k$. However, since 
high cardinality clusters occur only at low depths in the tree, the length of $Q$ for 
those iterations is relatively short.

After the first $\log_2{\frac{n}{2k}}$ iterations, we have that the cardinalities of clusters in $Q$ are at most $2k$, at which point, based on our assumption, we have that the length of $Q$ is close to $2d$.
